ISF Grant

### 2017-10-08 article.tex
references
references added

\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

@@

\subsection{Correctness properties}
Correctness property specifies the requirements for an implementation to be considered correct. Herlihy and Wing’s \emph{linearizability} property [*] is widely adopted for shared objects in conventional shared memory models.  Roughly speaking, linearizability requires that operations on objects appear to take effect instantaneously in some serial order, while preserving the happens before relation, i.e., if $op_1$ ends before $op_2$ begins, then $op_1$ should precede $op_2$ in the serial order. Linearizability has no notion of an aborted or failed operation, and requires that a process finish one operation before it invokes the next. The criteria according to which implementations that support crash-recovery failures are correct remains unclear.

Aguilera and Frølund have proposed \emph{strict linearizability} as a safety condition for persistent concurrent objects, which treats the crash of a process as a response, either successful or unsuccessful, to the interrupted operation [*]. A successful response means that the operation takes effect at some point between its invocation and the crash failure, and an unsuccessful response means that the operation does not take effect at all. Strict linearizability preserves both locality and program order. However, they proved that it precludes the implementation of some wait-free objects for certain machine models - there is no wait-free implementations of multi-reader single-writer (MRSW) registers from single-reader single-writer (SRSW) registers under strict linearizability.

Guerraoui and Levy have proposed \emph{persistent atomicity} [*]. It is similar to strict linearizability, but allows an operation interrupted by a failure to take effect before the subsequent invocation of the same process, possibly after the failure. They also proposed \emph{transient atomicity}, which relaxes this criterion even further and allows an interrupted operation to take effect before the subsequent write response of the same process. Both conditions ensure that the state of an object will be consistent in the wake of a crash, but it does not provide locality: correct histories of separate objects, when merged, will not necessarily yield a correct composite history.

Berryhill et al. have proposed an alternative, \emph{recoverable linearizability} [*], which achieves locality but may sacrifice program order after a crash. It is a relaxed version of persistent atomicity, which requires the operation to be linearizes or aborts before any subsequent linearization by the pending thread on that same object.

Izraelevitz et al. considered a real-world failure model, in which processes are assumed to fail together, as part of a full-system crash. Under this model, persistent atomicity and recoverable linearizability are indistinguishable (and thus local). The term \emph{durable linearizability} was used to refer to this merged safety condition under the restricted failure model.

@@

For this case, we present a different approach called \emph{recoverable response linearizability} (RR-linearizability) (good term????). In case of a failure, once the process is recovered it is "aware" of the failure, in the sense that it runs a designated recovery code. In a sense, this recovery code "extends" the interval of the interrupted operation. We require that the pending operation linearizes between it's invocation and the end of the recovery code, or aborts. In addition, in case of linearization, the returned value of the operation is known to the process. Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation is complete before the crash, but the response of it is lost due to the cache's lost, we allow the process, upon recovering, extending the interval of that operation using the recovery code (even though the operation was complete), such that at the end of it the process knows the returned value or aborts. This gives the process the flexibility to move the linearization point to after the crash, if needed. In addition, although the definition allows a process to abort the interrupted operation, it is equivalent to a definition in which a process must complete the operation - once a process knows the operation was aborted, it can reissue it.

@@

RR-linearizability does not specify the progress property required from the implementation. One can use a recoverable mutual exclusion lock at the recovery code in order to sequentially let the processes complete their pending operations. Obviously, this solution is not desired when the operation implementation we are recovering is wait-free. In this case, we would like the recovery code to be wait-free as well, meaning that a process finish it in a finite number of its own steps in the absence of failure. One more point that requires our attention is the possibility of a failure along the recovery code. Will it be realistic to assume that such a scenario will not happen? Of course, a recovery code that can suffer itself failures is better. However, this is not well defined. If we allow each failure to run new and different recovery code, we can use a nesting technique in a way which increase the running time with each failure. This may also requires infinite memory and code. A better approach will be such that in the case of failure inside of a recovery code, restarting the recovery code from some point is enough. In such case, if the recovery code is wait-free in the absence of a failure, we can guarantee that no matter how many failures the process experienced, once given enough time with no failure, it will finish the recovery code (and therefore the pending operation) in a wait-free manner.

The standard abstract model is consider in this work. The system consist of a set of asynchronous unreliable processes $P=\{p_1,\ldots,p_n\}$ that communicate by applying operations on shared \emph{base objects} that support atomic operations such as reads, writes, and read-modify-write primitives. All base objects are non-volatile, i.e, in case of a crash the values of the base objects is preserved. In order to implement a more complex object, such as stack and counter, each process is supplied with access procedures that simulate each operation on the implemented object using operations on base objects.

The failure model considered in this paper is individual. Each process may crash, recover, and then continue executing, at any point of the execution. Once a process crash, we assume that the system control knows at which line of the process code the failure occurs, or more precisely, at the course of which atomic operation the process failed. However, there is no induction whether the atomic operation took affect or not before the crash. Each process is also equipped with a recovery code for any possible failure, that is, for any point of the execution it specifies the recovery code in case of a crash at that point. Upon recovering, the system control moves the process to the beginning of the appropriate recovery code. In this sense, by running the recovery code, the process is aware of the crash. At the end of the recovery code, the process goes back to the main code, to the location specified by the recovery code. The above model can be implemented by storing the program counter at the non-volatile memory. In this case, after each operation on a base object, there should be an operation that updates the value of the program counter. This can be done by the system control, and the programmer need not be aware of it. Despite doubling the number of accesses to the main memory, this approach reduce the complexity of recovering in case of a crash.

Real world system presents more complications to the abstract model described above. In most modern system, each process has its own local volatile cache, which is used to store variables accessed by the process. Moreover, in order to improve performance in such systems, writes do not go directly to the shared memory, but rather stored in a buffer at the process's cache, while the system commit the writes later on according to the protocol and model it satisfies. Real system thus introduce reordering of instructions, as well as the possibility of losing large amount of data in case of a failure, which makes the recovery procedure harder and more complicate. However, the abstract model can be simulated in many of those architectures using special instructions as memory fences and avoid using the cache. The cost of such simulation nonetheless, might be very expensive in terms of running time.
The results presented in this paper implies that even in the simple abstract model the recovery procedure is not trivial. For future work, more complicated and realistic models needs to be studied.

@@

The correctness properties variants discussed earlier does not need to consider primitives, as any atomic operation satisfies any of the properties. Unlike it, RR-linearizability is not satisfies by primitives since the respond of the primitive is lost in case of a failure. Therefore, there is a need to supply each primitive with an implementation as well as a recovery code. A system which supply recoverable primitives can be used to implement any object in a recoverable way - in case of a crash the process will simply recover the last atomic operation it executed before the crash. In case the operation is aborted, the process will reissue it. After that, it can continue and execute the rest of its code safely. This observation focus our attention to implementation of primitives in a recoverable manner.

In the following section we present RR-liniarize implementations for well known primitives, as well as presenting impossibility result for others. We focus our attention on bounded wait-free implementations, that is, the number of steps a process takes when executing the recovery code in the absence of a failure is finite and bounded by a known constant (may be a function of n, the number of processes in the system), regardless of the other processes steps and the failures the process experienced before. In addition, we would like the recovery code to use a finite number of variables.
The fact that RR-linearizabilty allows us to swift the linearization point of an operation to after the crash is used to recover after a primitive failure.

@@

Test-and-Set (TAS) is a known object supported as a primitive in many modern architectures. A TAS object contains a binary value which initialized to 0, and support a single atomic operation TAS which set the value of the object to 1 and returns the old value.

We now show that there is no bounded wait-free RR-linearizability implementation for TAS for two processes, assuming the system supply only read/write and TAS primitives. We give a high level description for the proof. The construction also implies that any wait-free implementation need to use an unbounded number of TAS objects.

@@

Once $p$ is recover, it needs to execute a recovery code, since the value returned from $O_1$ is lost. We look at a solo run of $p$ starting from this point, and consider the possible return values. If $p$ returns 1, then we look at the execution where $q$ was the first to access $O_1$ before the crash. If $p$ returns 0 or abort, we look at the execution where $q$ is the second to access $O_1$. Notice that after the crash, $p$ is not able to distinguish between the different executions, thus it performs the same run after any of these. TAS can be seen as a consensus for two processes, the one that returns 1 wins. In any case we get a bivalet configuration - both processes want to return the same value (we consider abort as 0), while only one can return 1, while the other must return 0. Thus, it must be that both processes access another TAS object $O_2$. We can continue and use the same argument, constructing an execution where more and more different TAS objects are accessed.

\bibliographystyle{abbrv}
\bibliography{references}

>>>>

\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{cite}

@@

\subsection{Correctness properties} \label{Correctness properties}
Correctness property specifies the requirements for an implementation to be considered correct. Herlihy and Wing’s \emph{linearizability} property \cite{DBLP:journals/toplas/HerlihyW90} is widely adopted for shared objects in conventional shared memory models.  Roughly speaking, linearizability requires that operations on objects appear to take effect instantaneously in some serial order, while preserving the happens before relation, i.e., if $op_1$ ends before $op_2$ begins, then $op_1$ should precede $op_2$ in the serial order. Linearizability has no notion of an aborted or failed operation, and requires that a process finish one operation before it invokes the next. The criteria according to which implementations that support crash-recovery failures are correct remains unclear.

Aguilera and Fr{\o}lund have proposed \emph{strict linearizability} as a safety condition for persistent concurrent objects, which treats the crash of a process as a response, either successful or unsuccessful, to the interrupted operation \cite{Aguilera2003StrictLA}. A successful response means that the operation takes effect at some point between its invocation and the crash failure, and an unsuccessful response means that the operation does not take effect at all. Strict linearizability preserves both locality and program order. However, they proved that it precludes the implementation of some wait-free objects for certain machine models - there is no wait-free implementations of multi-reader single-writer (MRSW) registers from single-reader single-writer (SRSW) registers under strict linearizability.

Guerraoui and Levy have proposed \emph{persistent atomicity} \cite{DBLP:conf/icdcs/GuerraouiL04}. It is similar to strict linearizability, but allows an operation interrupted by a failure to take effect before the subsequent invocation of the same process, possibly after the failure. They also proposed \emph{transient atomicity}, which relaxes this criterion even further and allows an interrupted operation to take effect before the subsequent write response of the same process. Both conditions ensure that the state of an object will be consistent in the wake of a crash, but it does not provide locality: correct histories of separate objects, when merged, will not necessarily yield a correct composite history.

Berryhill et al. have proposed an alternative, \emph{recoverable linearizability} \cite{DBLP:conf/opodis/BerryhillGT15}, which achieves locality but may sacrifice program order after a crash. It is a relaxed version of persistent atomicity, which requires the operation to be linearizes or aborts before any subsequent linearization by the pending thread on that same object.

Izraelevitz et al. considered a real-world failure model, in which processes are assumed to fail together, as part of a full-system crash \cite{DBLP:conf/wdag/IzraelevitzMS16}. Under this model, persistent atomicity and recoverable linearizability are indistinguishable (and thus local). The term \emph{durable linearizability} was used to refer to this merged safety condition under the restricted failure model.

@@

For this case, we present a different approach called \emph{recoverable response linearizability} (RR-linearizability). In case of a failure, once the process is recovered it is "aware" of the failure, in the sense that it runs a designated recovery code. In a sense, this recovery code "extends" the interval of the interrupted operation. We require that the pending operation linearizes between it's invocation and the end of the recovery code, or aborts. In addition, in case of linearization, the returned value of the operation is known to the process. Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation took affect before the crash, but the response of it is lost due to the cache's lost, we allow the process, upon recovering, extending the interval of that operation using the recovery code, such that at the end of it the process knows the returned value or aborts. This gives the process the flexibility to move the linearization point to after the crash, if needed. In addition, although the definition allows a process to abort the interrupted operation, it is equivalent to a definition in which a process must complete the operation - once a process knows the operation was aborted, it can reissue it.

\iffalse
Progress properties are formally defined using histories. We follow [*] for these definitions. There are four types of steps: (1) an invocation step, denoted $(INV, p, X, op)$, represents the invocation by process $p$ of operation $op$ on implemented object $X$; (2) a response step, denoted $(RES, p, X, ret)$, represents the completion by process $p$ of the last operation it invoked on object $X$, with response $ret$; (3) a crash step, denoted $(CRASH, p)$, denotes a crash of process $p$; (4) a recovery step, denoted $(REC, p)$, denotes a recovery of process $p$;

A history $H$ is a sequence of steps, possibly involving multiple processes and implemented objects. For a given history $H$, we denote by $H \mid p$ the projection of $H$ onto the steps of process $p$. Similarly, we denote by $H \mid O$ the projection of $H$ onto the steps involving implemented object $O$. We adopt the convention that both $H \mid p$ and $H \mid O$ retain all crash steps in $H$. A response step is \emph{matching} with respect to an invocation step $s$ by a process $p$ on object $X$ in a history $H$ if it is the first response step by $p$ on $X$ that follows $s$ in $H$, and it occurs before p’s next invocation (if any) in $H$.
\fi

@@

RR-linearizability does not specify the progress property required from the implementation. One can use a recoverable mutual exclusion lock \cite{DBLP:conf/podc/GolabR16,DBLP:conf/podc/GolabH17} at the recovery code in order to sequentially let the processes complete their pending operations. Obviously, this solution is not desired when the operation implementation we are recovering is wait-free. In this case, we would like the recovery code to be wait-free as well, meaning that a process finish it in a finite number of its own steps in the absence of failure. Another point that requires our attention is the possibility of a failure along the recovery code. Will it be realistic to assume that such a scenario will not happen? Of course, a recovery code that can suffer itself failures is better. However, this is not well defined. If we allow each failure to run new and different recovery code, we can use a nesting technique in a way which increase the running time with each failure. This may also requires infinite memory and code. A better approach will be such that in the case of failure inside of a recovery code, restarting the recovery code from some point is enough. In such case, if the recovery code is wait-free in the absence of a failure, we can guarantee that no matter how many failures the process experienced, once given enough time with no failure, it will finish the recovery code (and therefore the pending operation) in a wait-free manner.

The standard abstract model is consider in this work. The system consist of a set of asynchronous unreliable processes $P=\{p_1,\ldots,p_n\}$ that communicate by applying operations on shared \emph{base objects} that support atomic operations such as reads, writes, and read-modify-write primitives. All base objects are non-volatile, i.e, in case of a crash the values stored in the base objects are preserved. In order to implement a more complex object, such as stack and counter, each process is supplied with access procedures that simulate each operation on the implemented object using operations on base objects.

The failure model considered in this paper is individual. Each process may crash, recover, and then continue executing, at any point of the execution. Once a process crash, we assume that the system control knows at which line of the code the failure occurs, or more precisely, at the course of which atomic operation the process failed. However, there is no induction whether the atomic operation took affect or not before the crash. Each process is also equipped with a recovery code for any possible failure, that is, for any point of the execution it specifies the recovery code in case of a crash at that point. Upon recovering, the system control moves the process to the beginning of the appropriate recovery code. In this sense, by running the recovery code, the process is aware of the crash. At the end of the recovery code, the process goes back to the main code, to the location specified by the recovery code. The above model can be implemented by storing the program counter at the non-volatile memory. In this case, after each operation on a base object, there should be an operation updating the value of the program counter. This can be done by the system control, and the programmer need not be aware of it. Despite doubling the number of accesses to the main memory, this approach reduce the complexity of recovering in case of a crash.

Real world system presents more complications to the abstract model described above. In most modern system, each process has its own local volatile cache, which is used to store variables accessed by the process. Moreover, in order to improve performance in such systems, writes do not go directly to the shared memory, but rather stored in a buffer at the process's cache, while the system commit the writes later on according to the protocol and model it satisfies. Real system thus introduce reordering of instructions, as well as the possibility of losing large amount of data in case of a failure, which makes the recovery procedure harder and more complicated. However, the abstract model can be simulated in many of those architectures, by using special instructions as memory fences and by avoid using the cache. The cost of such simulation, nonetheless, might be very expensive in terms of running time.
The results presented in this paper implies that even in the simple abstract model the recovery procedure is not trivial. For future work, more complicated and realistic models needs to be studied.

@@

The correctness properties variants discussed in section \ref{Correctness properties} does not need to consider primitives, as any atomic operation satisfies any of the properties. Unlike it, RR-linearizability is not satisfied by primitives since the respond of the primitive is lost in case of a failure. Therefore, there is a need to supply each primitive with an implementation as well as a recovery code. A system which supply recoverable primitives can be used to implement any object in a recoverable way - in case of a crash the process will simply recover the last atomic operation it executed before the crash. In case the operation is aborted, the process will reissue it. After that, it can continue and execute the remaining code safely. This observation focus our attention to implementation of primitives in a recoverable manner.

In the following section we present RR-linearizable implementations for well known primitives, as well as presenting impossibility result for others. We focus our attention on bounded wait-free implementations, that is, the number of steps a process takes when executing the recovery code in the absence of a failure is finite and bounded by a known constant (may be a function of n, the number of processes in the system), regardless of the other processes steps and the failures the process experienced so far. In addition, we would like the recovery code to use a finite number of variables.
The fact that RR-linearizabilty allows us to swift the linearization point of an operation to after the crash is used to recover after a primitive failure.

@@

Test-and-Set (TAS) is a known object supported as a primitive in many modern architectures. A TAS object contains a binary value initialized to 0, and support a single atomic operation TAS which set the value of the object to 1 and returns the old value.

We now show that there is no bounded wait-free RR-linearizable implementation for TAS even for two processes, assuming the system support only read, write and TAS primitives. We give a high level description for the proof. The construction also implies that any wait-free implementation need to use an unbounded number of TAS objects.

@@

Once $p$ is recover, it needs to execute a recovery code, since the value returned from $O_1$ is lost. We look at a solo run of $p$ starting from this point, and consider the possible return values. If $p$ returns 1, then we look at the execution where $q$ was the first to access $O_1$ before the crash. If $p$ returns 0 or abort, we look at the execution where $q$ is the second to access $O_1$. Notice that after the crash, $p$ does not able to distinguish between the different executions, thus it performs the same run after any of these. TAS can be seen as a consensus for two processes, the one that returns 1 wins. In any case we get a bivalet configuration - both processes want to return the same value (we consider abort as 0), while only one can return 1, while the other must return 0. Thus, it must be that both processes access another TAS object $O_2$. We can continue and use the same argument, constructing an execution where more and more different TAS objects are accessed.

\bibliographystyle{plain}
\bibliography{references} {}
